---
title: "Employees report: company 2"
author: "Write here your name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(readxl)
library(janitor)

i_am("report_company2.Rmd")

dt <- read_xlsx(here("data", "employees_company2.xlsx")) |>
    clean_names() |>
    group_by(department) |>
    mutate(salary = replace(salary, is.na(salary), median(salary, na.rm = TRUE))) |>
    ungroup()
```

## Overview

The company employs **`r nrow(dt)`** people.
The average salary is `r format(round(mean(dt$salary)), big.mark = ",")` dollars, and `r sum(dt$seniority == 0)` employees have just been hired.

## Salaries by department

```{r report-table}
report_table <- dt |>
    group_by(department) |>
    summarise(
        med_sal   = median(salary),
        total_n   = n(),
        new_hires = sum(seniority == 0)
    )

top_dept <- report_table |>
    filter(med_sal == max(med_sal)) |>
    pull(department)

knitr::kable(
    report_table,
    digits    = 0,
    col.names = c("Department", "Median salary", "Employees", "New hires")
)
```

```{r median-plot, fig.cap="Median salary by department"}
ggplot(report_table, aes(x = department, y = med_sal)) +
    geom_col() +
    labs(x = "", y = "Median salary ($)") +
    theme_minimal()
```

Key findings:

- The best paid department is **`r top_dept`**, with a median salary of `r format(round(max(report_table$med_sal)), big.mark = ",")` dollars.
- The largest department is `r report_table$department[which.max(report_table$total_n)]`, with `r max(report_table$total_n)` employees.
- Salaries grow with seniority in every department:

```{r seniority-plot, fig.cap="Salary and seniority by department"}
ggplot(dt, aes(x = seniority, y = salary, color = department)) +
    geom_point() +
    facet_wrap(~department) +
    labs(x = "Seniority (years since hiring)", y = "Salary ($)") +
    theme_minimal() +
    theme(legend.position = "none")
```

The analysis was carried out with the [tidyverse](https://www.tidyverse.org/).

## Data quality

```{r raw-data}
dt2 <- read_xlsx(here("data", "employees_company2.xlsx")) |>
    clean_names()

n_missing <- sum(is.na(dt2$salary))
```

The original file has `r n_missing` missing salaries.
We replaced each missing salary with the median salary of its department.

```{r imputation-table}
before <- dt2 |>
    group_by(department) |>
    summarise(missing = sum(is.na(salary)), mean_before = mean(salary, na.rm = TRUE))

after <- dt |>
    group_by(department) |>
    summarise(mean_after = mean(salary))

before |>
    left_join(after, by = "department") |>
    knitr::kable(
        digits    = 0,
        col.names = c("Department", "Missing salaries", "Mean before", "Mean after")
    )
```
