---
title: "Laboratory of Statistics and Mathematics - Data analysis report A"
format: html
author: "Write here your name"
---

```{r echo=FALSE}
# do not modify this!
knitr::opts_chunk$set(
    out.width = "50%", 
    fig.width = 10,     
    fig.asp = 9/16,   
    fig.align = "center" 
)
```

```{r dataset, echo=FALSE, warning=FALSE}
# do not modify this!
{
  seed <- length(rmarkdown::metadata$author)
  set.seed(seed)

  library(tidyverse) |> suppressPackageStartupMessages()
  n_users <- 1500
  cities <- c("Udine", "Padova", "Milano", "Trieste", "Treviso", "Torino")
  city <- sample(
    cities,
    size = n_users,
    replace = TRUE,
    prob = runif(cities)
  )

  device <- sample(
    c("mobile", "desktop"),
    size = n_users,
    replace = TRUE,
    prob = c(0.7, 0.3)
  )

  age <- round(rnorm(n_users, mean = 35, sd = 10))
  age <- pmin(pmax(age, 18), 70)  
  coupon_num <- rbinom(n_users, size = 1, prob = 0.3)
  coupon <- factor(coupon_num, levels = c(0, 1),
                  labels = c("no_coupon", "coupon"))
  city_effect_orders <- runif(length(cities), 0, 2)
  names(city_effect_orders) <- cities
  eff <- c(runif(1), -runif(1))
  device_effect_orders <- ifelse(device == "mobile", eff[1], eff[2])
  log_mu_orders <- log(1.4) +
    0.5 * coupon_num +                      
    device_effect_orders +
    unname(city_effect_orders[city])

  mu_orders <- exp(log_mu_orders)
  n_orders <- rnbinom(n_users, size=1, mu = mu_orders)
  spend_base <- rlnorm(n_users, meanlog = log(25), sdlog = 0.35)
  spend_per_order <- spend_base * if_else(coupon == "coupon", 0.9, 1.00)
  spend_per_order <- spend_per_order * (1 + (age - 35) / 1000)
  epsilon <- rnorm(n_users, mean = 0, sd = 0.10)  # random noise
  total_spent <- n_orders * spend_per_order * exp(epsilon)
  total_spent[n_orders == 0] <- 0
  avg_delivery_time <- rnorm(n_users, mean = 35, sd = 7) +
    ifelse(device == "mobile", -1, 1) 
  avg_delivery_time <- pmax(avg_delivery_time, 15)
  delivery_data <- tibble(
    user_id          = as.character(1:n_users),
    city             = city,
    device           = device,
    age              = age,
    coupon           = coupon,
    n_orders         = n_orders,
    total_spent      = round(total_spent, 2),
    avg_delivery_time = round(avg_delivery_time, 1)
  )

  user_data <- delivery_data |> select(user_id, age, city)
  user_data <- user_data[sample(1:n_users, size = n_users), ]
  delivery_data <- delivery_data |> select(-age, -city)
  delivery_data <- delivery_data[sample(1:n_users, size = n_users), ]

  dir.create("report_data")
  writexl::write_xlsx(delivery_data, "report_data/user_data.xlsx")
  writexl::write_xlsx(delivery_data, "report_data/delivery_data.xlsx")

}

```

# Instructions
## Intro
This `.rmd` file gives you access to two datasets called `delivery_data` and `user_data`, which gather information about a coupon-discount campaign run by a food-delivery company. Data are collected during a one-month window across different cities. A share of the user received a discount coupon at the beginning of the month. The company needs to understand if the campaign has been effective from a business perspective.

```{r glimpse}
head(user_data)
head(delivery_data)
```

Each row of the dataset represents a user. Columns denote:

- `user_id`: the unique identifier assigned to the user;
- `city`: city of reference for the specific user
- `device`: user's most used device;
- `age`: age of the user;
- `coupon`: denotes if the user has received the coupon or not;
- `n_orders`: number of orders placed by the user during the month;
- `total_spent`: total amount of money spent by the user during the month;
- `avg_delivery_time`: average delivery time to the user;

## Tasks
You are asked to complete/answer the following tasks/questions. When answering a question or completing a task, always show the code and interpret the results.

- Data exploration:
  1. Produce a table showing the number and percentage of users who received the coupon. 
  2. Produce a table showing the number and percentage of users who received the coupon by city.
  3. Does the share of users who placed at least one order changes according to wether the user received the coupon or not 
  4. Among under/over 35 y.o. users, plot the shares of users who placed at least one order based on wether the user received the coupon or not.
  5. Among users who placed at least one order, compute the mean and median of total expenditure for coupon and non-coupon users.
  6. Among users who placed at least one order, report the percentage of heavy users (more the five orders each) for coupon and non-coupon users.
- Inferential insights for business decisions:
  1. Compare the mean number of orders between coupon and non-coupon users. Report the results and interpret
  2. Compare the mean total expenditure between coupon and non-coupon users. Report the results and interpret
  3. Among users who received the coupon, does the mean per-user expenditure change between under/over 35 y.o.?
  4. Among users who placed at least one order, does the mean average order value differ between coupon and non-coupon users?
- Summarise your results. Is there evidence that the campaign has been effective?

**Hint:** for the inferential part, you are supposed to use the `t.test()` function to answer all questions.

To answer the questions, you can directly use this `.rmd` file by providing your answers in the [Solutions](#solutions) section. Alternatively, after running this `.rmd` once, you'll find the two datasets already available on you machine as `.xlsx` in a newly created `report_data` folder living in the same folder where this file is located. You can use them as you prefer to produce the report in `Word` or similar programs. Nevertheless, you will still have to answer all questions and provide your code.

# Solutions:

```{r}
library(tidyverse)
```

## Data exploration

## Inference 

