---
title: "Resources for Lab 2"

execute: 
  warning: false
  message: false
---

```{r}
#| echo: FALSE
knitr::opts_chunk$set(
    out.width = "50%", #
    fig.width = 4,     # reasonable size
    fig.asp = 0.618,   # golden ratio
    fig.align = "right" # mostly what I want
)
```
# Practice:

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(downloadthis)


library(randomNames)

set.seed(1)
n <- 150
depts <- c('Marketing', 'Accounting', 'Sales', 'Finance', 'HR', 'I.T.')

base_salary <- 50000

salary_depts <- tibble(
  department = depts, 
  dept_base_salary = rnorm(length(depts), mean = base_salary, sd = 5000),
  slope = runif(length(depts), 1000, 10000)  
)

empl <- tibble(full_name = randomNames(n)) |> 
  distinct() |> 
  separate(full_name, into = c("last_name", "first_name"), sep = ", ") |> 
  mutate(
    department = sample(depts, n, replace = TRUE),
    seniority = sample(0:15, size = n, replace = TRUE)
  ) |> 
  left_join(salary_depts, by = "department") |> 
  mutate(
    salary = dept_base_salary + slope * seniority + rnorm(n, 0, 3000),
    ID = 1000 + row_number()
  ) |> 
    select(-dept_base_salary, -slope)
writexl::write_xlsx(empl, "assets/class2/employees.xlsx")

empl|> 
    download_this(
    output_name = "employees",
    output_extension = ".xlsx",
    button_label = "Download employees.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```


::: {.cell .page-columns .page-full .column-screen .grey-section}


- Download the `employees.xlsx`.
- Create an R Project where to analyse the `employees.xlsx` dataset and move the `xlsx`file within to the project directory.
- Get the path to the data using the `here()` function.
- Import the dataset in R with the `read_xlsx()` function.

::: {.panel-tabset}
##### Exercise 1

- How many observations are collected in the dataset? Use the `dim()` function.

- What's the average salary in the company? Use either the function `mean()`or `summary()`

##### Exercise 2

- Use `filter()` and `arrange()` to show who are the three top earners in the `I.T` department.

- Use `group_by()`,  `summarise()` and `median()` to get the median salary in the `I.T.` department.

Construct a scatterplot for salary and seniority, colored by department. Store it in `my_plot`. Try adding a theme to the plot, for example:

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + theme_minimal()
```

What happens?

- Use `labs()` to modify the labels of the plot. For example:

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + theme_minimal() + labs(x="Seniority (years)")
```



##### Exercise 3

- How many employees have just been hired in the company?

- How many employees have just been hired in the I.T. department?

- Compute the average salary of employees from the Finance department with at least 8 years seniority.

##### Exercise 4

Compute a summary table where you report:

-  the median salary by department (using the function `median()`), 

- the number of employees by department (using the function `n()`), 

- the number of new hired by department.

Save it as an excel file via the `write_xlsx()` function.

##### Exercise 5

Given the summary table in the previous exercise, 
reproduce the following bar plot of the median salary by department (hint: use `geom_col()`).

```{r}
#| echo: FALSE
#| eval: TRUE
#| fig-align: right
#| fig-width: 4
#| fig-height: 2

empl |> 
    group_by(department) |> 
    summarise(median_salary=median(salary)) |> 
    ggplot(aes(x=department, y=median_salary)) +
    geom_col()
```

##### Exercise 6

You have to create an Excel file where employees from different department are collected in different sheets.

In R, you can create multi-sheets Excel files by feeding `write_xlsx()` with a list of datasets. Assume you stored the original dataset in the object `dt`. Now, try running

```{r}
#| echo: TRUE
#| eval: FALSE
my_departments <- split(dt, dt$department)
```

- What does the `split()`function do? Run `?split()` to read the documentation.

- What kind of object is stored in `my_departments`? Run `typeof(my_departments)` to investigate.

- Feed `write_xlsx()` with `my_departments`to create a multi-sheet Excel file.

:::
:::