---
title: "Resources for Lab 4"

execute: 
  warning: false
  message: false
---

```{r}
#| echo: FALSE
knitr::opts_chunk$set(
    out.width = "50%", #
    fig.width = 4,     # reasonable size
    fig.asp = 0.618,   # golden ratio
    fig.align = "right" # mostly what I want
)
```


::: {.cell .page-columns .page-full .column-screen .grey-section}



### Practice 1

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(ids)
library(downloadthis)

set.seed(123)
n_customers <- 200
n_purchases <- 500
unique_ids  <- uuid(n=n_customers)
unique_products <- c("laptop_1", "laptop_2", "tablet", "e-reader", "smartphone_1","smartphone_2", "earphones")
weights <- c(1,1,2,2,1,1,2)
prices <- c(2139.99, 899.99, 519.99, 199.99, 499.99, 299.99, 89.99)
prod_dict <- tibble(
    PRODUCT=unique_products, `Price (Euro)`=prices
)
prods_vec <- sample(unique_products, size=n_purchases, prob=weights, replace=TRUE)
ids_vec <- sample(unique_ids, size=n_purchases, prob = runif(n_customers), replace=TRUE)
purchases_history <- tibble(`customer id` = ids_vec, product=prods_vec)
my_list <- list(
    "purchases_history"=purchases_history,
    "products" = prod_dict
)


writexl::write_xlsx(my_list, "assets/class4/ecommerce.xlsx")

my_list|> 
    download_this(
    output_name = "ecommerce",
    output_extension = ".xlsx",
    button_label = "Download ecommerce.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```

- Download the `ecommerce.xlsx` file and move it in the project folder. If you open it in Excel, you'll see it is made of two sheets, named "purchases_history" and "products".

- You will need to import both sheets in `R`. 

For importing, use the `read_xlsx()`function and specify the `sheet` argument.
```{r}
#| echo: TRUE
#| eval: FALSE

# declare where you are in your project folder
i_am("practice1.R")

# construct the path to your data
path <- here("data", "ecommerce.xlsx")

# read the two sheets by specifying the "sheet" argument
purchases <- read_xlsx(path = path, sheet = "purchases_history")
products  <- read_xlsx(path = path, sheet = "products")
```


::: {.panel-tabset}

#### Exercise 1

- Clean the names of the two data frames by usign `clean_names()`from the `janitor` package.
- Compute the number of sales by product

#### Exercise 2

Compute the total amount of sales by product in thousands of euros.

:::
:::


### Practice 2

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(ids)
library(downloadthis)
library(moderndive)
set.seed(123)

dt <- avocados |> 
  select(date, type, average_price) |> 
  group_by(date, type) |> 
  summarise(avg_price = mean(average_price), .groups = "drop") |> 
  pivot_wider(names_from = type, values_from = avg_price)

dt |> 
    download_this(
    output_name = "avocados",
    output_extension = ".xlsx",
    button_label = "Download avocados.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
writexl::write_xlsx(dt, "assets/class4/avocados.xlsx")

# dt |> 
#   pivot_longer(cols=c(conventional, organic), names_to = "type", values_to = "av_price") |> 
#   ggplot(aes(x=date, y=av_price, col=type))+
#   geom_line()

```

- Download the `avocados.xlsx` file and move it in your project folder. 

- The dataset contains the average prices ($) of conventional and organic avocados, collected from 2015-01-04 to 2018-03-25. 

::: {.panel-tabset}
#### Exercise 1

Transform the data to match

```{r}
#| echo: FALSE

dt |> 
  pivot_longer(cols=c(conventional, organic), names_to = "type", values_to = "avg_price") 

```

and then use `ggplot` and `geom_line()` to plot the average price trajectory across the years by type of avocado
:::

### Practice 3

Organise the answers to [Lab 3 - Practice 1](class3.qmd) in a [R markdown](https://rmarkdown.rstudio.com/lesson-2.html) file.