---
title: "Resources for Lab 2"

execute:
  warning: false
  message: false
---

```{r}
#| echo: FALSE
knitr::opts_chunk$set(
    out.width = "50%", #
    fig.width = 4,     # reasonable size
    fig.asp = 0.618,   # golden ratio
    fig.align = "right" # mostly what I want
)
```


::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice 1: Import & summarise

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(downloadthis)


library(randomNames)

set.seed(1)
n <- 150
depts <- c('Marketing', 'Accounting', 'Sales', 'Finance', 'HR', 'I.T.')

base_salary <- 50000

salary_depts <- tibble(
  department = depts,
  dept_base_salary = rnorm(length(depts), mean = base_salary, sd = 5000),
  slope = runif(length(depts), 1000, 10000)
)

empl <- tibble(full_name = randomNames(n)) |>
  distinct() |>
  separate(full_name, into = c("last_name", "first_name"), sep = ", ") |>
  mutate(
    department = sample(depts, n, replace = TRUE),
    seniority = sample(0:15, size = n, replace = TRUE)
  ) |>
  left_join(salary_depts, by = "department") |>
  mutate(
    salary = dept_base_salary + slope * seniority + rnorm(n, 0, 3000),
    ID = 1000 + row_number()
  ) |>
    select(-dept_base_salary, -slope)
writexl::write_xlsx(empl, "assets/class2/employees.xlsx")

empl|>
    download_this(
    output_name = "employees",
    output_extension = ".xlsx",
    button_label = "Download employees.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```


- Download the `employees.xlsx`.
- Create an R Project where to analyse the `employees.xlsx` dataset and move the `xlsx` file within the project directory.
- Get the path to the data using the `here()` function.
- Import the dataset in R with the `read_xlsx()` function.

::: {.panel-tabset}
##### Exercise 1

- How many observations are collected in the dataset? Use the `dim()` function.

- What's the average salary in the company? Use either the function `mean()`or `summary()`

##### Exercise 2

- Use `filter()` and `arrange()` to check who are the three top earners in the `I.T` department.

- Use `group_by()`,  `summarise()` and `median()` to get the median salary in the `I.T.` department.

Using `geom_point()`, construct a scatterplot for salary and seniority, colored by department. Store it in `my_plot`. Try adding a theme to the plot, for example:

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + theme_minimal()
```

What happens?

- Use `labs()` to modify the labels of the plot. For example:

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + theme_minimal() + labs(x="Seniority (years)", y="Salary", color="Department:")
```

- Use `facet_grid()` to create a multi-window plot

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + facet_grid(~department)
```

- Export your plot with the function `ggsave()`

```{r}
#| echo: TRUE
#| eval: FALSE
my_great_plot <- my_plot + theme_minimal() + labs(x="Seniority (years)", y="Salary", color="Department:")+ facet_grid(~department)

ggsave(my_great_plot, filename = here("great_plot.jpeg"))

```

##### Exercise 3

- Use `group_by()` and `summarise()` to compute the average and total expenditure for salaries by department

##### Exercise 4

- How many employees have just been hired in the company?

- How many employees have just been hired in the I.T. department?

- Compute the average salary of employees from the Finance department with at least 8 years seniority.

##### Exercise 5

Create a summary table where you report:

- the median salary by department (using the function `median()`),

- the number of employees by department (using the function `n()`),

- the number of new hired by department.

Save it as an excel file via the `write_xlsx()` function.

##### Exercise 6

Given the summary table in the previous exercise,
reproduce the following bar plot of the median salary by department (hint: use `geom_col()`).

```{r}
#| echo: FALSE
#| eval: TRUE
#| fig-align: right
#| fig-width: 4
#| fig-height: 2

empl |>
    group_by(department) |>
    summarise(median_salary=median(salary)) |>
    ggplot(aes(x=department, y=median_salary)) +
    geom_col()
```

##### Exercise 7

You have to create an Excel file where employees from different department are collected in different sheets.

In R, you can create multi-sheets Excel files by feeding `write_xlsx()` with a list of datasets. Assume you stored the original `employees.xlsx`dataset in the dataframe `dt`. Now, try running

```{r}
#| echo: TRUE
#| eval: FALSE
my_departments <- split(dt, dt$department)
```

- What does the `split()`function do? Run `?split()` to read the documentation.

- What kind of object is stored in `my_departments`? Run `typeof(my_departments)` to investigate.

- Feed `write_xlsx()` with `my_departments`to create a multi-sheet Excel file.

:::
:::


::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice 2: Cleaning

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(downloadthis)


library(randomNames)

set.seed(3)
n <- 200
depts <- c('Marketing', 'Accounting', 'Sales', 'Finance', 'HR', 'I.T.')

base_salary <- 50000

salary_depts <- tibble(
  department = depts,
  dept_base_salary = rnorm(length(depts), mean = base_salary, sd = 5000),
  slope = runif(length(depts), 1000, 10000)
)

empl <- tibble(full_name = randomNames(n)) |>
  distinct() |>
  separate(full_name, into = c("last_name", "first_name"), sep = ", ") |>
  mutate(
    department = sample(depts, n, replace = TRUE),
    seniority = sample(0:15, size = n, replace = TRUE)
  ) |>
  left_join(salary_depts, by = "department") |>
  mutate(
    salary = dept_base_salary + slope * seniority + rnorm(n, 0, 3000),
    ID = 1000 + row_number()
  ) |>
    select(-dept_base_salary, -slope) |>
    mutate(salary=map_dbl(salary, ~if_else(rbinom(1,1,.9)==0, NA, .x))) |>
    rename(`Last NAME`=last_name, `First-name`=first_name, `Seniority`=seniority)
writexl::write_xlsx(empl, "assets/class3/raw_data_employees.xlsx")

empl|>
    download_this(
    output_name = "employees_company2",
    output_extension = ".xlsx",
    button_label = "Download employees_company2.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```

- The HR department sent you the employees file from a different company.
- Download the `employees_company2.xlsx` and import it in R.
- As you can see, the file has a structure that is very similar to the previous database. Nevertheless, it presents two critical issues:
    - Column names do not follow a consistent notation. Sometimes they are capitalized, sometimes they are not. Some times they includes spaces, sometimes they do not.
    - There are missing values

::: {.panel-tabset}
##### Exercise 1

Use the function `clean_names()` from the `janitor` package to tidy up column names automatically

##### Exercise 2

- How many missing values are present in the salary column? Use the `summary()` function.
- Replace missing values with the median salary of department the missing value belongs to.

:::
:::


::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice 3: Joining data frames

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(ids)
library(downloadthis)

set.seed(123)
n_customers <- 200
n_purchases <- 500
unique_ids  <- uuid(n=n_customers)
unique_products <- c("laptop_1", "laptop_2", "tablet", "e-reader", "smartphone_1","smartphone_2", "earphones")
weights <- c(1,1,2,2,1,1,2)
prices <- c(2139.99, 899.99, 519.99, 199.99, 499.99, 299.99, 89.99)
prod_dict <- tibble(
    PRODUCT=unique_products, `Price (Euro)`=prices
)
prods_vec <- sample(unique_products, size=n_purchases, prob=weights, replace=TRUE)
ids_vec <- sample(unique_ids, size=n_purchases, prob = runif(n_customers), replace=TRUE)
purchases_history <- tibble(`customer id` = ids_vec, product=prods_vec)
my_list <- list(
    "purchases_history"=purchases_history,
    "products" = prod_dict
)


writexl::write_xlsx(my_list, "assets/class4/ecommerce.xlsx")

my_list|>
    download_this(
    output_name = "ecommerce",
    output_extension = ".xlsx",
    button_label = "Download ecommerce.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```

- Download the `ecommerce.xlsx` file and move it in the project folder. If you open it in Excel, you'll see it is made of two sheets, named "purchases_history" and "products".

- You will need to import both sheets in `R`.

For importing, use the `read_xlsx()`function and specify the `sheet` argument.
```{r}
#| echo: TRUE
#| eval: FALSE

# declare where you are in your project folder
i_am("practice3.R")

# construct the path to your data
path <- here("data", "ecommerce.xlsx")

# read the two sheets by specifying the "sheet" argument
purchases <- read_xlsx(path = path, sheet = "purchases_history")
products  <- read_xlsx(path = path, sheet = "products")
```


::: {.panel-tabset}

#### Exercise 1

- Clean the names of the two data frames by usign `clean_names()`from the `janitor` package.
- Compute the number of sales by product

#### Exercise 2

Compute the total amount of sales by product in thousands of euros.

:::
:::


::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice 4: Pivoting

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(ids)
library(downloadthis)
library(moderndive)
set.seed(123)

dt <- avocados |>
  select(date, type, average_price) |>
  group_by(date, type) |>
  summarise(avg_price = mean(average_price), .groups = "drop") |>
  pivot_wider(names_from = type, values_from = avg_price)

dt |>
    download_this(
    output_name = "avocados",
    output_extension = ".xlsx",
    button_label = "Download avocados.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
writexl::write_xlsx(dt, "assets/class4/avocados.xlsx")
```

- Download the `avocados.xlsx` file and move it in your project folder.

- The dataset contains the average prices ($) of conventional and organic avocados, collected from 2015-01-04 to 2018-03-25.

::: {.panel-tabset}
#### Exercise 1

Transform the data to match

```{r}
#| echo: FALSE

dt |>
  pivot_longer(cols=c(conventional, organic), names_to = "type", values_to = "avg_price")

```

and then use `ggplot` and `geom_line()` to plot the average price trajectory across the years by type of avocado
:::
:::
