---
title: "Resources for Lab 3"

execute: 
  warning: false
  message: false
---

```{r}
#| echo: FALSE
knitr::opts_chunk$set(
    out.width = "50%", #
    fig.width = 4,     # reasonable size
    fig.asp = 0.618,   # golden ratio
    fig.align = "right" # mostly what I want
)
```



::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice 1

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(downloadthis)


library(randomNames)

set.seed(1)
n <- 150
depts <- c('Marketing', 'Accounting', 'Sales', 'Finance', 'HR', 'I.T.')

base_salary <- 50000

salary_depts <- tibble(
  department = depts, 
  dept_base_salary = rnorm(length(depts), mean = base_salary, sd = 5000),
  slope = runif(length(depts), 1000, 10000)  
)

empl <- tibble(full_name = randomNames(n)) |> 
  distinct() |> 
  separate(full_name, into = c("last_name", "first_name"), sep = ", ") |> 
  mutate(
    department = sample(depts, n, replace = TRUE),
    seniority = sample(0:15, size = n, replace = TRUE)
  ) |> 
  left_join(salary_depts, by = "department") |> 
  mutate(
    salary = dept_base_salary + slope * seniority + rnorm(n, 0, 3000),
    ID = 1000 + row_number()
  ) |> 
    select(-dept_base_salary, -slope)
writexl::write_xlsx(empl, "assets/class2/employees.xlsx")

empl|> 
    download_this(
    output_name = "employees",
    output_extension = ".xlsx",
    button_label = "Download employees.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```


- Download the `employees.xlsx`.
- Create an R Project where to analyse the `employees.xlsx` dataset and move the `xlsx` file within the project directory.
- Get the path to the data using the `here()` function.
- Import the dataset in R with the `read_xlsx()` function.

::: {.panel-tabset}
##### Exercise 1

- How many observations are collected in the dataset? Use the `dim()` function.

- What's the average salary in the company? Use either the function `mean()`or `summary()`

##### Exercise 2

- Use `filter()` and `arrange()` to check who are the three top earners in the `I.T` department.

- Use `group_by()`,  `summarise()` and `median()` to get the median salary in the `I.T.` department.

Using `geom_point()`, construct a scatterplot for salary and seniority, colored by department. Store it in `my_plot`. Try adding a theme to the plot, for example:

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + theme_minimal()
```

What happens?

- Use `labs()` to modify the labels of the plot. For example:

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + theme_minimal() + labs(x="Seniority (years)", y="Salary", color="Department:")
```

- Use `facet_grid()` to create a multi-window plot

```{r}
#| echo: TRUE
#| eval: FALSE
my_plot + facet_grid(~deprtment)
```


##### Exercise 3

- Use `group_by()` and `summarise()` to compute the average and total expenditure for salaries by department

##### Exercise 4

- How many employees have just been hired in the company?

- How many employees have just been hired in the I.T. department?

- Compute the average salary of employees from the Finance department with at least 8 years seniority.

##### Exercise 5

Create a summary table where you report:

- the median salary by department (using the function `median()`), 

- the number of employees by department (using the function `n()`), 

- the number of new hired by department.

Save it as an excel file via the `write_xlsx()` function.

##### Exercise 6

Given the summary table in the previous exercise, 
reproduce the following bar plot of the median salary by department (hint: use `geom_col()`).

```{r}
#| echo: FALSE
#| eval: TRUE
#| fig-align: right
#| fig-width: 4
#| fig-height: 2

empl |> 
    group_by(department) |> 
    summarise(median_salary=median(salary)) |> 
    ggplot(aes(x=department, y=median_salary)) +
    geom_col()
```

##### Exercise 7

You have to create an Excel file where employees from different department are collected in different sheets.

In R, you can create multi-sheets Excel files by feeding `write_xlsx()` with a list of datasets. Assume you stored the original `employees.xlsx`dataset in the dataframe `dt`. Now, try running

```{r}
#| echo: TRUE
#| eval: FALSE
my_departments <- split(dt, dt$department)
```

- What does the `split()`function do? Run `?split()` to read the documentation.

- What kind of object is stored in `my_departments`? Run `typeof(my_departments)` to investigate.

- Feed `write_xlsx()` with `my_departments`to create a multi-sheet Excel file.

:::
:::


::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice 2

```{r}
#| echo: FALSE
#| eval: TRUE
library(tidyverse)
library(downloadthis)


library(randomNames)

set.seed(3)
n <- 200
depts <- c('Marketing', 'Accounting', 'Sales', 'Finance', 'HR', 'I.T.')

base_salary <- 50000

salary_depts <- tibble(
  department = depts, 
  dept_base_salary = rnorm(length(depts), mean = base_salary, sd = 5000),
  slope = runif(length(depts), 1000, 10000)  
)

empl <- tibble(full_name = randomNames(n)) |> 
  distinct() |> 
  separate(full_name, into = c("last_name", "first_name"), sep = ", ") |> 
  mutate(
    department = sample(depts, n, replace = TRUE),
    seniority = sample(0:15, size = n, replace = TRUE)
  ) |> 
  left_join(salary_depts, by = "department") |> 
  mutate(
    salary = dept_base_salary + slope * seniority + rnorm(n, 0, 3000),
    ID = 1000 + row_number()
  ) |> 
    select(-dept_base_salary, -slope) |> 
    mutate(salary=map_dbl(salary, ~if_else(rbinom(1,1,.9)==0, NA, .x))) |> 
    rename(`Last NAME`=last_name, `First-name`=first_name, `Seniority`=seniority)
writexl::write_xlsx(empl, "assets/class3/raw_data_employees.xlsx")

empl|> 
    download_this(
    output_name = "employees_company2",
    output_extension = ".xlsx",
    button_label = "Download employees_company2.xlsx",
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
```

- The HR department sent you the employees file from a different company.
- Download the `employees_company2.xlsx` and import it in R.
- As you can see, the file has a structure that is very similar to the previous database. Nevertheless, it presents two critical issues:
    - Column names do not follow a consistent notation. Sometimes they are capitalized, sometimes they are not. Some times they includes spaces, sometimes they do not. 
    - There are missing values

::: {.panel-tabset}
##### Exercise 1

Use the function `clean_names()` from the `janitor` package to tidy up column names automatically

##### Exercise 2

- How many missing values are present in the salary column? Use the `summary()` function.
- Replace missing values with the median salary of department the missing value belongs to.

:::
:::