---
title: "Resources for Lab 3"

execute:
  warning: false
  message: false
---

::: {.cell .page-columns .page-full .column-screen .grey-section}

### Practice: the employees report

In Lab 2 you analysed the employees of two companies with R scripts. Now you will turn that analysis into a **report**: a single `.Rmd` file that mixes text, code and results, and that you can re-knit whenever the data change.

You need the two datasets from Lab 2. If you no longer have them, download them again:

```{r}
#| echo: FALSE
library(downloadthis)

download_file(
  path = "assets/class2/employees.xlsx",
  button_label = "Download employees.xlsx",
  button_type = "success",
  has_icon = TRUE,
  icon = "fa fa-save",
  class = "button_small"
)

download_file(
  path = "assets/class3/raw_data_employees.xlsx",
  output_name = "employees_company2",
  button_label = "Download employees_company2.xlsx",
  button_type = "success",
  has_icon = TRUE,
  icon = "fa fa-save",
  class = "button_small"
)
```

**Before you start**

- Open the R Project you used in Lab 2: double-click its `.Rproj` file, or use *File > Open Project...* in RStudio. Check that the project name appears in the top-right corner of RStudio.
- Put both Excel files in the `data` folder of the project, renaming the second one `employees_company2.xlsx` if needed. Your project should look like this:

```
my_project/
├── my_project.Rproj
├── report.Rmd          <- you will create this now
└── data/
    ├── employees.xlsx
    └── employees_company2.xlsx
```

- Go to *File > New File > R Markdown...*. In the window that opens, choose *Document*, set the title to "Employees report", write your name as author, select *HTML* and click *OK*. If RStudio asks to install some packages, accept.
- RStudio opens an example document. Save it with *File > Save* as `report.Rmd`, directly in the project folder (not inside `data`).
- Click **Knit**. An HTML page with the example content should open, and a `report.html` file appears next to `report.Rmd`.
- Now look at the example: the YAML header between `---`, the text, the grey code chunks. Then delete everything **below** the first code chunk (the one starting with ```` ```{r setup, include=FALSE} ````).

::: {.panel-tabset}
##### Exercise 1

Setup.

- Add a table of contents to the output by editing the YAML header. Replace the line `output: html_document` with the following (indentation matters: use 2 spaces per level):

```yaml
output:
  html_document:
    toc: true
```

- Complete the `setup` chunk so that it loads the packages, declares where the report is and imports the data. Replace each `___` with the right value:

````markdown
```{{r setup, include=FALSE}}
knitr::opts_chunk$set(echo = TRUE)

# packages
library(tidyverse)
library(here)
library(readxl)
library(janitor)

# declare where this file is within the project
i_am("___")

# import the data
dt <- read_xlsx(here("___", "___"))
```
````

- Before knitting, run the chunk with the green ▶ button in its top-right corner, then type `dt` in the console. Do you see the data? Knit only when the chunk works.
- The `setup` chunk has the option `include=FALSE`. Knit the document: do you see the code or the data anywhere? What does this option do?
- Change `echo = TRUE` into `echo = FALSE, message = FALSE` in `knitr::opts_chunk$set()`. You will see the effect in the next exercises: the code of all the following chunks will be hidden.

::: {.callout-tip collapse="true"}
## Common errors when knitting

- `` `path` does not exist ``: the file is not where you said it is. Check that `employees.xlsx` is inside the `data` folder (or in whatever folder matching your `here()` call), that the name is spelled exactly the same (capital letters matter), and that `report.Rmd` is in the project folder.
- `Could not find associated project in working directory`: the file name in `i_am()` does not match the name of your `.Rmd`, or you are not working inside the R Project.
- `object 'dt' not found`: knitting starts from an empty session and ignores what you ran in the console. Every object the report uses must be created by a chunk in the `.Rmd`, above the chunk that uses it.
- `there is no package called ...`: install the missing package from the console with `install.packages()`. Never put `install.packages()` inside the `.Rmd`.
:::

##### Exercise 2

Write a section called `## Overview` with a short paragraph reporting:

- the number of employees in the company,
- their average salary,
- the number of new hires (employees with `seniority` equal to 0).

Do not type any number by hand: use inline R code, for example

````markdown
The company employs `r knitr::inline_expr("nrow(dt)")` people.
````

What happens to the average salary if you just use `round(mean(dt$salary))`? Try `format()` with the `big.mark` argument.

##### Exercise 3

Write a section called `## Salaries by department`.

- Add a chunk that computes the summary table of Lab 2 (median salary, number of employees and number of new hires by department) and prints it with `knitr::kable()`. Use the `digits` and `col.names` arguments to make it readable.
- Add a chunk with the bar plot of the median salary by department. Give it a caption with the chunk option `fig.cap`:

````markdown
```{{r median-plot, fig.cap="Median salary by department"}}
...
```
````

##### Exercise 4

Below the plot, write a bullet list of key findings.

- One item should state which department has the highest median salary, in **bold**. Compute the name in a chunk and store it in an object (for example `top_dept`), then use it with inline code.
- Add the salary vs seniority scatterplot, faceted by department, from Lab 2.
- Add a link to the [tidyverse website](https://www.tidyverse.org/).

##### Exercise 5

HR now wants the same report for company 2.

- In the *Files* pane of RStudio, tick `report.Rmd` and use *More > Copy...* to make a copy called `report_company2.Rmd`. Open the copy, change the title in the YAML header, and change `i_am("report.Rmd")` into `i_am("report_company2.Rmd")`.
- First, try the naive way: in the `setup` chunk of the copy, replace `employees.xlsx` with `employees_company2.xlsx` and knit. Why does it fail? (Look at the column names of the two files.)
- Fix it: in the `setup` chunk, clean the column names of the company 2 data with `clean_names()` and replace the missing salaries with the median salary of their department, as you did in [Lab 2](sol2.qmd#exercise-2-1). Store the result in `dt` and knit again.
- Open `report.html` and `report_company2.html` side by side. You now have two complete reports, one per company, but you only wrote the text once: every number, table and plot was updated by R. That's why we write reports in R Markdown!

##### Exercise 6

The readers of the company 2 report should know that some salaries were imputed. In `report_company2.Rmd`, write a section called `## Data quality`.

- The `setup` chunk already replaced the missing salaries in `dt`. To count them, add a new chunk that reads the raw file again into `dt2`, cleaning the names but without imputing. You don't need `i_am()` again: it is already in the `setup` chunk.

```{r}
#| eval: FALSE
dt2 <- read_xlsx(here("___", "___")) |>
    clean_names()
```

- In a sentence, report how many salaries are missing (inline code again). Hint: `is.na()` returns `TRUE` for each missing value, and `sum()` counts the `TRUE`s.
- Show a table with, for each department, the number of missing salaries, the average salary before the imputation (from `dt2`, with `na.rm = TRUE`) and after it (from `dt`). Hint: compute two summary tables by department and join them with `left_join()`.

##### Exercise 7

- Change the output of one of the reports to `word_document` and knit. Compare the result with the HTML version.

:::
:::
