---
title: "Homework"
---

Two data analysis reports, each worth up to **1 additional point** on the final exam (2 points in total). Each homework generates **your own dataset** from your student ID, so every student works on different data.

```{r}
#| echo: FALSE
htmltools::tagList(lapply(c("homework_A.Rmd", "homework_B.Rmd"), \(f)
  downloadthis::download_file(
    path = file.path("assets/homework", f),
    button_label = paste("Download", f),
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
))
```

| Homework | Topic | Deadline |
|---|---|---|
| A | Effectiveness of a discount-coupon campaign | TBD |
| B | Performance of online ad campaigns | TBD |

## Instructions

1. Create a new R Project for each homework, and move the downloaded `.Rmd` file into the project folder.
2. Open the `.Rmd`, write your name in the `author` field of the YAML header, and your student ID (matricola) in the first chunk: `student_id <- 123456`.
3. Knit the document. The data are created by the chunks at the top of the file, and the instructions and questions follow.
4. Write your answers in the **Solutions** section, at the end of the document. **Do not modify the chunks above it**: they generate your data.
5. Knit to PDF and check the result.

## Knitting to PDF

The homework must be knitted to PDF, which requires LaTeX. Install it once, from the R console:

```{r}
#| eval: FALSE
install.packages("tinytex")
tinytex::install_tinytex()
```

Then restart RStudio. The first knit can take a while, because missing LaTeX packages are downloaded automatically.

::: {.callout-tip}
## Tips

- Code lines are not wrapped in the PDF: keep them short, and break long function calls over several lines.
- If you really cannot install LaTeX, change the output to `word_document`, knit, and save the Word file as PDF.
:::

## Submission

Send both the `.Rmd` and the `.pdf` files to *giuseppe.alfonzetti@uniud.it* by the deadline.

## Evaluation

Each homework is evaluated from 0 to 1 point:

- **Code and reporting (0.5 points)**: the `.Rmd` knits without errors; the code is correct and readable; tables and plots are clear, with meaningful labels; numbers in the text come from inline R code, not typed by hand.
- **Interpretation and insights (0.5 points)**: every result is explained in words; the conclusions answer the business questions and are supported by the numbers.
