---
title: "Solutions Lab 3 practice"

execute:
  warning: false
  message: false
---

You can download the complete solutions and knit them in your Lab 2 project (the data must be in the `data` folder):

```{r}
#| echo: FALSE
htmltools::tagList(lapply(c("report.Rmd", "report_company2.Rmd"), \(f)
  downloadthis::download_file(
    path = file.path("assets/class3", f),
    button_label = paste("Download", f),
    button_type = "success",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_small"
  )
))
```

Below, we go through it exercise by exercise.

## Exercise 1

The YAML header sets the title, author, date and output format. `toc: true` adds a table of contents, and `toc_float: true` keeps it visible on the side while you scroll. The date is inline R code, so it updates every time you knit.

````markdown
---
title: "Employees report"
author: "Write here your name"
output:
  html_document:
    toc: true
    toc_float: true
---
````

The `setup` chunk runs all the code the rest of the report needs. With `include=FALSE`, the chunk is run but neither its code nor its output appears in the report. `knitr::opts_chunk$set()` sets default options for all the following chunks. Here, `echo = FALSE` hides the code, because the reader of a report wants results, not R code.

````markdown
```{{r setup, include=FALSE}}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(readxl)
library(janitor)

i_am("report.Rmd")

dt <- read_xlsx(here("data", "employees.xlsx"))
```
````

Recall that when you knit, R starts from a clean session. Objects in your environment are not visible, so everything the report uses must be created inside the `.Rmd`.

## Exercise 2

Every number is computed with inline code:

````markdown
## Overview

The company employs **`r knitr::inline_expr("nrow(dt)")`** people.
The average salary is `r knitr::inline_expr('format(round(mean(dt$salary)), big.mark = ",")')` dollars, and `r knitr::inline_expr("sum(dt$seniority == 0)")` employees have just been hired.
````

When printing inline, knitr writes large numbers in scientific notation, so `round(mean(dt$salary))` shows up as something like `1.04 × 10⁵`. `format()` with `big.mark = ","` prints `103,789` instead.

## Exercise 3

`knitr::kable()` turns a data frame into a formatted table. `digits` rounds numbers and `col.names` replaces the column names in the output only (the data frame is unchanged).

````markdown
```{{r report-table}}
report_table <- dt |>
    group_by(department) |>
    summarise(
        med_sal   = median(salary),
        total_n   = n(),
        new_hires = sum(seniority == 0)
    )

knitr::kable(
    report_table,
    digits    = 0,
    col.names = c("Department", "Median salary", "Employees", "New hires")
)
```

```{{r median-plot, fig.cap="Median salary by department"}}
ggplot(report_table, aes(x = department, y = med_sal)) +
    geom_col() +
    labs(x = "", y = "Median salary ($)") +
    theme_minimal()
```
````

## Exercise 4

We add the name of the best paid department to the `report-table` chunk, so that it is available in the text:

```{r}
#| eval: FALSE
top_dept <- report_table |>
    filter(med_sal == max(med_sal)) |>
    pull(department)
```

`pull()` extracts a column as a plain vector. Then we write the findings as a markdown bullet list, with the scatterplot in between:

````markdown
Key findings:

- The best paid department is **`r knitr::inline_expr("top_dept")`**, with a median salary of `r knitr::inline_expr('format(round(max(report_table$med_sal)), big.mark = ",")')` dollars.
- The largest department is `r knitr::inline_expr("report_table$department[which.max(report_table$total_n)]")`, with `r knitr::inline_expr("max(report_table$total_n)")` employees.
- Salaries grow with seniority in every department:

```{{r seniority-plot, fig.cap="Salary and seniority by department"}}
ggplot(dt, aes(x = seniority, y = salary, color = department)) +
    geom_point() +
    facet_wrap(~department) +
    labs(x = "Seniority (years since hiring)", y = "Salary ($)") +
    theme_minimal() +
    theme(legend.position = "none")
```

The analysis was carried out with the [tidyverse](https://www.tidyverse.org/).
````

## Exercise 5

After copying `report.Rmd` into `report_company2.Rmd`, update the title and the `i_am()` call so that they refer to the new file:

```{r}
#| eval: FALSE
i_am("report_company2.Rmd")
```

If you only change the file name of the data, knitting stops with `object 'seniority' not found`. The raw company 2 file calls that column `Seniority`, and it also has missing salaries, which would break the medians. The data must be cleaned and imputed first, so in the `setup` chunk replace the line reading `employees.xlsx` with

```{r}
#| eval: FALSE
dt <- read_xlsx(here("data", "employees_company2.xlsx")) |>
    clean_names() |>
    group_by(department) |>
    mutate(salary = replace(salary, is.na(salary), median(salary, na.rm = TRUE))) |>
    ungroup()
```

The cleaning code is the one from Lab 2. The only new part is `ungroup()`, which removes the grouping so it does not affect later computations.

Knit again. You now have `report.html` and `report_company2.html`: two complete reports, one per company, while you wrote the text only once.

## Exercise 6

In `report_company2.Rmd`, `dt` has no missing values anymore. To count them, we read the raw file again, cleaning the names only:

````markdown
## Data quality

```{{r raw-data}}
dt2 <- read_xlsx(here("data", "employees_company2.xlsx")) |>
    clean_names()

n_missing <- sum(is.na(dt2$salary))
```

The original file has `r knitr::inline_expr("n_missing")` missing salaries.
We replaced each missing salary with the median salary of its department.
````

For the before/after table, we summarise the raw data (`dt2`) and the imputed data (`dt`) separately, and join them by department:

```{r}
#| eval: FALSE
before <- dt2 |>
    group_by(department) |>
    summarise(missing = sum(is.na(salary)), mean_before = mean(salary, na.rm = TRUE))

after <- dt |>
    group_by(department) |>
    summarise(mean_after = mean(salary))

before |>
    left_join(after, by = "department") |>
    knitr::kable(
        digits    = 0,
        col.names = c("Department", "Missing salaries", "Mean before", "Mean after")
    )
```

## Exercise 7

To get a Word document, change the output in the YAML header to

```yaml
output: word_document
```

## The complete files

### `report.Rmd`

```{r}
#| echo: FALSE
#| comment: ""
#| class-output: markdown
cat(readLines("assets/class3/report.Rmd"), sep = "\n")
```

### `report_company2.Rmd`

```{r}
#| echo: FALSE
#| comment: ""
#| class-output: markdown
cat(readLines("assets/class3/report_company2.Rmd"), sep = "\n")
```
