---
title: "Solutions Lab 4 practice"

execute: 
  warning: false
  message: false
---

## Practice 1

For importing, use the `read_xlsx()`function and specify the `sheet` argument.
```{r}
#| echo: TRUE
#| eval: FALSE
library(readxl)
library(tidyverse)

# declare where you are in your project folder
i_am("practice1.R")

# construct the path to your data
path <- here("data", "ecommerce.xlsx")

# read the two sheets by specifying the "sheet" argument
purchases <- read_xlsx(path = path, sheet = "purchases_history")
products  <- read_xlsx(path = path, sheet = "products")
```

```{r}
#| echo: FALSE
#| eval: TRUE
library(readxl)
library(tidyverse)

path <- "assets/class4/ecommerce.xlsx"
purchases <- read_xlsx(path = path, sheet = "purchases_history")
products  <- read_xlsx(path = path, sheet = "products")

```

### Exercise 1


```{r}
library(janitor)

purchases <- purchases |> clean_names()
products <- products |> clean_names()
```

To compute the total number of sales by product we can use the `group_by()` + `summarise()` combo. 

```{r}
purchases |> 
    group_by(product) |> 
    summarise(
        number_of_sales = n()
    )
```

There is even a faster way to count how many times the values of a given variable appear in the dataset. Literally, you can pipe the `count()` function

```{r}
purchases |> 
    count(product)
```


### Exercise 2

To compute the total amount of money spent on each product we first join the two datasets, 

```{r}
purchases |> 
    left_join(products, by = "product") 
```

and then comput the totals,

```{r}
purchases |> 
    left_join(products, by = "product") |> 
    group_by(product) |> 
    summarise(
        total = sum(price_euro),
        total_thousands = sum(price_euro)/1000
        )
```

Equivalently, we can also first count how many items have been sold by product type, as in the previous exercise, and then join the resulting table with the price listings


```{r}
purchases |> 
    count(product) |> 
    left_join(products)
```

Once here, we need to compute total by product by multiplying their prices and quantities sold


```{r}
purchases |> 
    count(product) |> 
    left_join(products) |> 
    mutate(
        total_thousands = price_euro * n /1000
    )
```

## Practice 2
### Exercise 1

```{r}
#| echo: TRUE
#| eval: FALSE
library(readxl)
library(tidyverse)

# declare where you are in your project folder
i_am("practice2.R")

# construct the path to your data
path <- here("data", "avocados.xlsx")

avocados <- read_xlsx(path = path)
```

```{r}
#| echo: FALSE
#| eval: TRUE
library(readxl)
library(tidyverse)

path <- "assets/class4/avocados.xlsx"
avocados <- read_xlsx(path = path)

```

We need to use `pivot_longer()` to reshape the data:

```{r}
new_avo <- avocados |> 
  pivot_longer(
    cols = c(conventional, organic), 
    names_to = "type", 
    values_to = "avg_price"
    ) 
new_avo
```

Once reshaped, we can plot the price trajectories over time:

```{r}
new_avo |> 
  ggplot(aes(x=date, y=avg_price, col=type))+
  geom_line()
```

As always, we can customise the plots as we prefer:

```{r}
library(ggokabeito)
new_avo |> 
  ggplot(aes(x=date, y=avg_price, col=type))+
  geom_line(linewidth=1) +
  labs(x="", y= "Average price ($)", color="Avocado origin:")+
  theme_bw() +
  theme(
    legend.position = "top", 
    panel.grid.minor = element_blank()
    )+
  scale_color_okabe_ito()
```