---
title: "Data manipulation"
subtitle: "Laboratory of Statistics and Mathematics 2025/2026"
author: Giuseppe Alfonzetti
format: 
  revealjs:
    slide-number: true
    incremental: false   
    footer: "Data manipulation"
    theme: ../themes/slides_theme.scss
    highlight-style: arrow-light
    transition: convex
    transition-speed: fast
title-slide-attributes:
    data-background-color: "#1565C0"
    data-footer: " "
embed-resources: true
highlight-style: tango

---

## R basics - filtering {style="font-size: 0.5em;"}

Consider the `data.frame`
```{r}
#| echo: FALSE
#| eval: TRUE
var0 <- 1:5
var1 <- c("pop-up", "banner", "video", "banner", "pop-up")
var2 <- c(56, 23, 321, 89, 10)
var3 <- c(7,2,25, 10, 23)
var4 <- c(TRUE, FALSE, FALSE, TRUE, TRUE)
my_df <- data.frame(
    "id" = var0,
    "ad_type"= var1, 
    "n_clicks"=var2, 
    "sales"=var3, 
    "weekend"=var4)
```
```{r}
#| echo: TRUE
#| eval: TRUE
my_df
```

. . .

**Create a filter:**

```{r}
#| echo: TRUE
#| eval: TRUE
my_filter = my_df$sales > 20
my_filter
```

. . . 

**View:**

```{r}
#| echo: TRUE
#| eval: TRUE
my_df[my_filter, ]
```

. . .

**Overwrite:**

```{r}
#| echo: TRUE
#| eval: TRUE
my_df[my_filter, ]$ad_type = c("banner", "banner")
my_df
```

## R basics - Missing values {style="font-size: 0.5em;"}

Consider the `data.frame`

```{r}
var0 <- 1:5
var1 <- c("pop-up", "banner", "video", "banner", "pop-up")
var2 <- c(56, 23, 321, NA, 10)
var3 <- c(7,NA,25, 10, 23)
var4 <- c(TRUE, FALSE, FALSE, TRUE, TRUE)
my_df <- data.frame(
    "id" = var0,
    "ad_type"= var1, 
    "n_clicks"=var2, 
    "sales"=var3, 
    "weekend"=var4)
```

```{r}
#| echo: TRUE
#| eval: TRUE
my_df
```

. . .

```{r}
#| echo: TRUE
#| eval: TRUE
summary(my_df)
```


## R basics - Missing values {style="font-size: 0.5em;"}

Consider the `data.frame`

```{r}
var0 <- 1:5
var1 <- c("pop-up", "banner", "video", "banner", "pop-up")
var2 <- c(56, 23, 321, NA, 10)
var3 <- c(7,NA,25, 10, 23)
var4 <- c(TRUE, FALSE, FALSE, TRUE, TRUE)
my_df <- data.frame(
    "id" = var0,
    "ad_type"= var1, 
    "n_clicks"=var2, 
    "sales"=var3, 
    "weekend"=var4)
```

```{r}
#| echo: TRUE
#| eval: TRUE
my_df
```


**Locate:**

```{r}
#| echo: TRUE
#| eval: TRUE
clicks_missing <- is.na(my_df$n_clicks)
clicks_missing
```

. . .

**Replace:**

```{r}
#| echo: TRUE
#| eval: TRUE
my_df$n_clicks[clicks_missing] <- median(my_df$n_clicks, na.rm=TRUE)
my_df
```


## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

```{r}
library(palmerpenguins)
```

Consider this dataset:

```{r}
#| echo: TRUE
#| eval: TRUE
library(tidyverse)
penguins
```

**Select:**

```{r}
#| echo: TRUE
#| eval: TRUE
my_penguins <- select(penguins, species, island, sex, bill_length_mm, body_mass_g, year)
my_penguins
```

## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

**Filter:**

```{r}
#| echo: TRUE
#| eval: TRUE
filter(my_penguins, sex=="female")
```

. . .

```{r}
#| echo: TRUE
#| eval: TRUE
filter(my_penguins, sex=="female", island=="Biscoe")
```

## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

**Mutate:**

```{r}
#| echo: TRUE
#| eval: TRUE
mutate(my_penguins, body_mass_kg=body_mass_g/1000)
```

. . .

```{r}
#| echo: TRUE
#| eval: TRUE
mutate(my_penguins, body_mass_g=body_mass_g+1)
```

## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

**Mutate according to logical conditions:**

```{r}
#| echo: TRUE
#| eval: TRUE
my_penguins
```

. . . 

```{r}
#| echo: TRUE
#| eval: TRUE
mutate(
    my_penguins, 
    bill_length_mm = replace(bill_length_mm, bill_length_mm>39, 100)
)
```


## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

**Replace Missing values**

```{r}
#| echo: TRUE
#| eval: TRUE
my_penguins
```

. . .

```{r}
#| echo: TRUE
#| eval: TRUE
mutate(my_penguins, body_mass_g_filled = replace(body_mass_g, is.na(body_mass_g), median(body_mass_g, na.rm=TRUE)))
```

## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

**Replace Missing values**

```{r}
#| echo: TRUE
#| eval: TRUE
mutate(
    my_penguins, 
    body_mass_g = replace(body_mass_g, is.na(body_mass_g), median(body_mass_g, na.rm=TRUE)),
    bill_length_mm = replace(bill_length_mm, is.na(bill_length_mm), median(bill_length_mm, na.rm=TRUE))
)
```

## Data manipulation with `tidyverse` {style="font-size: 0.5em;"}

**Pipes:**

```{r}
#| echo: TRUE
#| eval: TRUE
penguins
```

. . .

```{r}
#| echo: TRUE
#| eval: FALSE
#| code-line-numbers: "1|2|3|4|5|6"
penguins |> 
    select(species, island, sex, body_mass_g, bill_length_mm) |> 
    mutate(body_mass_kg=body_mass_g/1000) |> 
    filter(sex=="female") |> 
    arrange(bill_length_mm) |> 
    select(-body_mass_g)
```

. . .

```{r}
#| echo: FALSE
#| eval: TRUE
penguins |> 
    select(species, island, sex, body_mass_g, bill_length_mm) |> 
    mutate(body_mass_kg=body_mass_g/1000) |> 
    filter(sex=="female") |> 
    arrange(bill_length_mm)|> 
    select(-body_mass_g)
```


## Groupwise manipulation {style="font-size: 0.5em;"}

**Summarise by group:**
```{r}
#| echo: TRUE
#| eval: FALSE
#| code-line-numbers: "1|2|3"
penguins |> 
    group_by(species) |> 
    summarise(average_bill_length = mean(bill_length_mm))
```

. . .

```{r}
#| echo: FALSE
#| eval: TRUE
penguins |> 
    group_by(species) |> 
    summarise(average_bill_length = mean(bill_length_mm))
```

. . .

```{r}
#| echo: TRUE
#| eval: FALSE
#| code-line-numbers: "3"
penguins |> 
    group_by(species) |> 
    summarise(average_bill_length = mean(bill_length_mm, na.rm=TRUE))
```

. . .

```{r}
#| echo: FALSE
#| eval: TRUE
penguins |> 
    group_by(species) |> 
    summarise(average_bill_length = mean(bill_length_mm, na.rm=TRUE))
```


## Groupwise manipulation {style="font-size: 0.5em;"}

**Mutate by group:**
```{r}
#| echo: TRUE
#| eval: FALSE
#| code-line-numbers: "1|2|3|4"
penguins |> 
    select(species, island, bill_length_mm) |> 
    group_by(species) |> 
    mutate(species_average_bill_length = mean(bill_length_mm, na.rm=TRUE))
```

. . .

```{r}
#| echo: FALSE
#| eval: TRUE
penguins |> 
    select(species, island, bill_length_mm) |> 
    group_by(species) |> 
    mutate(species_average_bill_length = mean(bill_length_mm, na.rm=TRUE))
```

. . .

```{r}
#| echo: TRUE
#| eval: FALSE
#| code-line-numbers: "5|6"
penguins |> 
    select(species, island, bill_length_mm) |> 
    group_by(species) |> 
    mutate(
        species_average_bill_length = mean(bill_length_mm, na.rm=TRUE),
        dist_from_species_average = bill_length_mm - species_average_bill_length
        )
```

. . .

```{r}
#| echo: FALSE
#| eval: TRUE
penguins |> 
    select(species, island, bill_length_mm) |> 
    group_by(species) |> 
    mutate(
        species_average_bill_length = mean(bill_length_mm, na.rm=TRUE),
        dist_from_species_average = bill_length_mm - species_average_bill_length
        )
```


## Save objects {style="font-size: 0.5em;"}


```{r}
#| echo: TRUE
#| eval: TRUE
my_res <- penguins |> 
    group_by(species) |> 
    summarise(average_bill_length = mean(bill_length_mm, na.rm=TRUE))
my_res
```

**Store as an `R` data object:**

```{r}
#| echo: TRUE
#| eval: FALSE
# To save the object in your current project folder under the name "my_results.rda"
save(my_res, here("my_results.rda"))

# To load the object when you need it next time
load(here("my_results.rda"))
```

. . .

**Store as an `Excel` file:**

```{r}
#| echo: TRUE
#| eval: FALSE

# To save the object in a spreadsheet in your current project folder under the name "my_results.rda"
write_xlsx(my_res, path = here("my_results.xlsx"))
```

