---
title: "Comparing means"
subtitle: "Laboratory of Statistics and Mathematics 2025/2026"
author: Giuseppe Alfonzetti
format: 
  revealjs:
    slide-number: true
    incremental: false   
    footer: "Comparing means - the t-test"
    theme: ../themes/slides_theme.scss
    highlight-style: arrow-light
    transition: convex
    transition-speed: fast
title-slide-attributes:
    data-background-color: "#1565C0"
    data-footer: " "
embed-resources: true
highlight-style: tango

---

```{r}
library(tidyverse)
```

# Types of t-tests {style="font-size: 0.7em;"}

- One-sample t-test : Compare one mean vs a known/target value

. . . 

- Two-sample t-test: Compare means of two independent groups

. . .

- Paired t-test: Compare means on paired data (e.g. same people, before and after treatment)

# Case study A {style="font-size: 0.7em;"}

. . .

- Your company is running a **marketing campaign** and you are monitoring the performance of the landing page in terms of *time-on-page* (i.e. seconds spent by the user on the landing page) to **assess its effectiveness**. 

. . .

- Your direct competitor reached a time-on-page equal to $10$ seconds.

. . .

- You want to assess if your campaign performance is **significantly different** from your competitor benchmark.

# Data {style="font-size: 0.7em;"}

```{r}
set.seed(123)
n_obs <- 300
landing_data <- tibble(
  version = rep(c("standard_campaign", "modified_campaign"), each = n_obs/2),
  time_on_page = c(
      rnorm(n_obs/2, mean = 9.8, sd = 1.0),  
      rnorm(n_obs/2, mean = 12, sd = 2.0)   
    )
)
landing_data <- landing_data[sample(1:n_obs, n_obs),] |> 
  mutate(
    user_id = paste0("user_", 1:n_obs)
  ) |> 
  select(user_id, everything())

modified_landing_data <- landing_data |> 
  filter(version=="modified_campaign")

landing_data <- landing_data |> 
  filter(version=="standard_campaign")
```

```{r}
#| echo=TRUE
landing_data
```

# Plot {style="font-size: 0.7em;"}
```{r}
#| echo=TRUE
landing_data |>
  ggplot(aes(x = version, y = time_on_page)) +
    geom_point() +
    labs(x="", y="Time on page (s)")
```

# Plot {style="font-size: 0.7em;"}

```{r}
#| echo: TRUE
#| code-line-numbers: "4"
landing_data |>
  ggplot(aes(x = version, y = time_on_page)) +
    geom_point() +
    stat_summary(fun = "mean", geom = "point", size = 4, color="red")+
    labs(x="", y="Time on page (s)")
```

# Plot {style="font-size: 0.7em;"}

```{r}
#| echo: TRUE
#| code-line-numbers: "3"
landing_data |>
  ggplot(aes(x = version, y = time_on_page)) +
    geom_boxplot() +
    labs(x="", y="Time on page (s)")
```

# Business question {style="font-size: 0.7em;"}

::: {.columns}
::: {.column width="50%"}
Is the marketing campaign performing significantly different from the competitor benchmark, in terms of time-on-page?
:::
::: {.column width="50%"}
```{r}
#| echo=TRUE
landing_data |>
  group_by(version) |>
  summarise(
    n = n(),
    mean_time = mean(time_on_page),
    sd_time = sd(time_on_page)
  )
```
:::
:::

. . .

We can compare the average time-on-page to the benchmark with a t-test, using the `t.test()`function

. . . 

```{r}
#| echo: TRUE
#| code-line-numbers: "1,5|2|3|4"
t.test(
  x=landing_data$time_on_page, 
  mu=10, 
  alternative = "two.sided"
  )
```

# Case study B {style="font-size: 0.7em;"}

- You assessed that the data highlight a small underperformance of your company campaign compared to the competitor one. 

. . .

- The design team provided you with an alternative version of the landing page. 

. . .

- You are interested to check if the modified landing page performs differently from the old one.

# Case study B - Experiments {style="font-size: 0.7em;"}

Now you can choose between:

. . . 

- Bring to the new landing pege a *new sample of users* to monitor the **time-on-page**;

. . .

- Bring to the new landing page the *same users* who visited the old one. 

. . .

In the first case, you can compare the average time-on-page of the standard and modified landing pages using a **Two sample t-test**. In the second case, you have to rely on a **Paired t-test**.

# Data {style="font-size: 0.7em;"}

```{r}
#| echo: TRUE
modified_landing_data
```

. . .

```{r}
#| echo: TRUE
full_data <- landing_data |> 
  bind_rows(modified_landing_data)
full_data
```


# Plot two groups {style="font-size: 0.7em;"}

```{r}
#| echo=TRUE
full_data |>
  ggplot(aes(x = version, y = time_on_page)) +
    geom_boxplot() +
    labs(x="", y="Time to event (s)")
```

# Business question {style="font-size: 0.7em;"}

::: {.columns}
::: {.column width="50%"}
Is the modified landing page performing significantly different from the old one, in terms of time-on-page?
:::
::: {.column width="50%"}
```{r}
#| echo=TRUE
full_data |>
  group_by(version) |>
  summarise(
    n = n(),
    mean_time = mean(time_on_page),
    sd_time = sd(time_on_page)
  )
```
:::
:::

. . . 

We can compare the average time-on-page between standard and modified landing pages, using the `t.test()`function

```{r}
#| echo: TRUE
#| eval: FALSE
t.test(time_on_page ~ version, data = full_data, paired=FALSE)
```

. . .

```{r}
#| echo: FALSE
t.test(time_on_page ~ version, data = full_data)
```

