---
title: "Multiple groups comparison"
subtitle: "Laboratory of Statistics and Mathematics 2025/2026"
author: Giuseppe Alfonzetti
format: 
  revealjs:
    slide-number: true
    incremental: false   
    footer: "ANOVA"
    # theme: ../themes/slides_theme.scss
    highlight-style: arrow-light
    transition: convex
    transition-speed: fast
title-slide-attributes:
    data-background-color: "#1565C0"
    data-footer: " "
embed-resources: true
highlight-style: tango

---

## Case Study {style="font-size: 0.7em;"}

A retail chain asks:

> **Do customers from different cities spend differently in our stores?**

- *Different cities* (Milan, Paris, Madrid, Berlin);
- Total *purchase amount* by customer.

. . .

**Goal:**

- Help the retail chain understanding **in which city** customers spend more on average.

---

## ANOVA tests {style="font-size: 0.7em;"}

Compares averages of **3+ groups**.

> Does average spending differ across **cities**?

. . .

*Intuition: the ANOVA test compares*

- **Within-city variation**;
- **Between-city variation** .

. . .

If *between-city variation* is "larger" than *within-city variation*, then group averages differ!

. . . 

Statistical Hypoteses:

- $H_0$: all means are equal;
- $H_1$: at least one mean differs;


## Case A {style="font-size: 0.6em;"}


```{r}
library(ggplot2)
library(tidyverse)
set.seed(123)
dfA <- tibble(
  city = rep(c("Milan", "Berlin", "Paris", "Madrid"), each = 40),
  spend = round(c(
            rnorm(40, 50, 20),
            rnorm(40, 50, 30),
            rnorm(40, 60, 10),
            rnorm(40, 55, 10)),2)
)|> 
    mutate(
  spend = if_else(spend < 10, 10, spend))
```

::: {.columns}
::: {.column width="40%"}
```{r}
#| echo: TRUE
dfA
```

```{r}
#| echo: TRUE
dfA |> 
    group_by(city) |> 
    summarise(
        av_spend = mean(spend),
        sd_spend = sd(spend),
        n = n()
        )
```

:::
::: {.column width="60%"}


```{r}
#| echo: TRUE
dfA |> 
    ggplot(aes(x = city, y = spend, fill = city))+
    geom_boxplot()+
    theme_bw() +
    theme(legend.position = "none")+
    labs(x="", y="Total expenditure by customer (€)")
```

```{r}
#| echo: TRUE
anova_modelA <- aov(spend~city, data=dfA)
summary(anova_modelA)
```

:::


:::

## Case B {style="font-size: 0.6em;"}


```{r}
library(ggplot2)
library(tidyverse)
set.seed(123)
dfB <- tibble(
  city = rep(c("Milan", "Berlin", "Paris", "Madrid"), each = 40),
  spend = round(c(
            rnorm(40, 50, 3),
            rnorm(40, 50, 4),
            rnorm(40, 60, 5),
            rnorm(40, 55, 5)),2)
) |> 
    mutate(
  spend = if_else(spend < 10, 10, spend))
```

::: {.columns}
::: {.column width="40%"}
```{r}
#| echo: TRUE
dfB
```

```{r}
#| echo: TRUE
dfB |> 
    group_by(city) |> 
    summarise(
        av_spend = mean(spend),
        sd_spend = sd(spend),
        n = n()
        )
```

:::
::: {.column width="60%"}
```{r}
#| echo: TRUE
dfB |> 
    ggplot(aes(x = city, y = spend, fill = city))+
    geom_boxplot()+
    theme_bw() +
    theme(legend.position = "none")+
    labs(x="", y="Total expenditure by customer (€)")
```

```{r}
#| echo: TRUE
anova_modelB <- aov(spend~city, data=dfB)
summary(anova_modelB)
```

:::
:::

## Which city differs? {style="font-size: 0.7em;"}

. . .

```{r}
#| echo: TRUE
TukeyHSD(anova_modelA)
```

. . .

```{r}
#| echo: TRUE
TukeyHSD(anova_modelB)
```


## A special case of `lm()` {style="font-size: 0.6em;"}

::: {.columns}
::: {.column width="50%"}
```{r}
#| echo: TRUE
summary(anova_modelA)
```
:::
::: {.column width="50%"}
```{r}
#| echo: TRUE
lm_modelA <- lm(spend ~ city, data = dfA)
summary(lm_modelA)
anova(lm_modelA)
```
:::



