---
title: "Laboratory of Statistics and Mathematics - Data analysis report A"
author: "Write here your name"
output:
  pdf_document:
    toc: true
---

```{r student-id}
# write your student ID (matricola) here, e.g. student_id <- 123456
student_id <- NA
```

```{r setup, include=FALSE}
# do not modify this!
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    out.width = "80%",
    fig.width = 7,
    fig.asp = 0.6,
    fig.align = "center"
)
```

```{r dataset, echo=FALSE}
# do not modify this!
if (is.na(as.integer(student_id))) stop("Write your student ID in the first chunk of the document.")
set.seed(as.integer(student_id))

library(tidyverse)

hw_data <- local({
  n_users <- 1500

  cities    <- c("Milano", "Torino", "Padova", "Udine", "Treviso", "Trieste")
  city_size <- c(0.30, 0.22, 0.15, 0.12, 0.11, 0.10) * exp(rnorm(6, 0, 0.1))
  city_eff  <- c(0.20, 0.10, 0, 0, -0.05, -0.05) + rnorm(6, 0, 0.1)
  city_time <- c(3, 2, 0, -2, -1, -2) + rnorm(6, 0, 1)
  names(city_eff) <- names(city_time) <- cities
  city <- sample(cities, size = n_users, replace = TRUE, prob = city_size)

  device <- sample(c("mobile", "desktop"), size = n_users, replace = TRUE, prob = c(0.7, 0.3))
  age <- round(rnorm(n_users, mean = 35, sd = 10))
  age <- pmin(pmax(age, 18), 70)

  coupon_num <- rbinom(n_users, size = 1, prob = 0.3)
  coupon <- factor(coupon_num, levels = c(0, 1), labels = c("no_coupon", "coupon"))

  avg_delivery_time <- rnorm(n_users, mean = 35, sd = 7) +
    unname(city_time[city]) +
    ifelse(device == "mobile", -1, 1)
  avg_delivery_time <- pmax(avg_delivery_time, 15)

  coupon_eff <- pmax(0.6 - 0.03 * (age - 35), 0.05) * runif(1, 0.8, 1.2)
  log_mu_orders <- log(1.5) +
    coupon_num * coupon_eff +
    ifelse(device == "mobile", runif(1, 0.2, 0.5), 0) -
    0.015 * (age - 35) -
    0.02 * (avg_delivery_time - 35) +
    unname(city_eff[city])
  n_orders <- rnbinom(n_users, size = 1, mu = exp(log_mu_orders))

  spend_per_order <- rlnorm(n_users, meanlog = log(25), sdlog = 0.35) *
    (1 + 0.005 * (age - 35)) *
    if_else(coupon == "coupon", 0.9, 1)
  total_spent <- n_orders * spend_per_order * exp(rnorm(n_users, 0, 0.1))
  total_spent[n_orders == 0] <- 0

  delivery_data <- tibble(
    user_id           = as.character(1:n_users),
    city              = city,
    device            = device,
    age               = age,
    coupon            = coupon,
    n_orders          = n_orders,
    total_spent       = round(total_spent, 2),
    avg_delivery_time = round(avg_delivery_time, 1)
  )

  user_data <- delivery_data |> select(user_id, age, city)
  user_data <- user_data[sample(1:n_users, size = n_users), ]
  delivery_data <- delivery_data |> select(-age, -city)
  delivery_data <- delivery_data[sample(1:n_users, size = n_users), ]

  list(user_data = user_data, delivery_data = delivery_data)
})

user_data <- hw_data$user_data
delivery_data <- hw_data$delivery_data
rm(hw_data)
```

# Instructions

## Intro

A food-delivery company ran a discount-coupon campaign. At the beginning of the month, a share of its users received a coupon giving a discount on their orders. During the month, the company collected data about its users across different cities. The company needs to understand whether the campaign has been effective from a business perspective.

The data are stored in two data frames, `user_data` and `delivery_data`, which are already available in your environment once you run the chunks above:

```{r glimpse}
head(user_data)
head(delivery_data)
```

Each row of `user_data` describes a user:

- `user_id`: the unique identifier of the user;
- `age`: age of the user;
- `city`: city of the user.

Each row of `delivery_data` describes the activity of a user during the month:

- `user_id`: the unique identifier of the user;
- `device`: the device most used by the user;
- `coupon`: whether the user received the coupon or not;
- `n_orders`: number of orders placed by the user during the month;
- `total_spent`: total amount of money (euros) spent by the user during the month;
- `avg_delivery_time`: average delivery time (minutes) of the user's orders.

## Tasks

Answer the following questions in the [Solutions](#solutions) section. For each question:

- show the code you used;
- present your results with tables (`knitr::kable()`) and/or plots, whichever you find more appropriate;
- **interpret the results in words**, reporting the relevant numbers with inline R code (do not type numbers by hand).

1. Join the two data frames into a single one. How many users does it contain? Check that no user has been lost.
2. Produce a table with the number and percentage of users who received the coupon, overall and by city.
3. Compute the share of users who placed at least one order, for coupon and non-coupon users. Show it with a bar plot.
4. Split the users into two age groups (under 35 and 35 or older). Does the difference found in question 3 hold in both groups? Show it with a plot.
5. Focus on the users who placed at least one order. For coupon and non-coupon users, compute the mean and median of the total expenditure and of the average order value (total expenditure divided by the number of orders). Use a plot to compare the distributions of the average order value.
6. Among the users who placed at least one order, compute the percentage of heavy users (more than five orders), for coupon and non-coupon users.
7. Does the picture change between mobile and desktop users? Produce a table with one row per device, the share of users with at least one order among coupon and non-coupon users in two separate columns, and their difference in a third column (hint: `pivot_wider()`).
8. Summarise your results. The coupon lowers the price of each order, but it may increase the number of orders: compare the average revenue per user between coupon and non-coupon users. Would you recommend running the campaign again?

**Do not modify the chunks above this section.** Knit the document to PDF and submit both the `.Rmd` and the `.pdf` files.

# Solutions

```{r}
library(tidyverse)
```

## Question 1

## Question 2

## Question 3

## Question 4

## Question 5

## Question 6

## Question 7

## Question 8
